---
title: "V1 LD score regression on immune mediated GWAS"
output: html_notebook
---

# LD score regression on auotoimmunity GWAS
## Step 1: prepare the datasets

The tutorial and info on the package and how to run the code are here:  
https://github.com/GenomicSEM/GenomicSEM/wiki/3.-Models-without-Individual-SNP-effects

```{r Libraries}
library(GenomicSEM)
library(data.table)
library(pheatmap)
library(R.utils)
library(stringr)
library(tidyr)
library(dplyr)
library(corrplot)
```


```{r Load the files}
crohn_delange <- fread('Summary_Stats/delange-2017_cd_build37_40266_20161107.txt', 
                       data.table = F)
ulccol_delange <- fread('Summary_Stats/delange-2017_uc_build37_45975_20161107.txt', 
                        data.table = F)
allergies_ferreira <- fread('Summary_Stats/ferreira-2017_allergies_build37_SHARE-without23andMe.LDSCORE-GC.SE-META.v0', 
                            data.table = F)
alzheimer_kunkle <- fread('Summary_Stats/kunkle-2019_alzheimer_build37_Kunkle_etal_Stage1_results.txt',
                          data.table = F)
ssc_lopez <- fread('Summary_Stats/lopez-2019_ssc_build37_Lopez-Isac_prePMID_META_GWAS_SSc.meta.txt',
                   data.table = F)
ra_okada <- fread('Summary_Stats/okada-2014_RA_build37_RA_GWASmeta_TransEthnic_v2.txt.gz.gz', 
                  data.table = F)
asthma_demeanis <- fread('Summary_Stats/demeanis-2017_asthma_build37_TAGC_Multiancestry_and_European-Ancestry_Meta-analyses_Results.tsv',
                         data.table = F)
alzheimer_wightman <- fread('Summary_Stats/wightman-2021_alzheimer_build37_PGCALZ2sumstatsExcluding23andMe.txt.gz',
                            data.table = F)
```

### Add rsID

Load the reference SNP to add the rsIDs to the GWAS that do not have them.
```{r}
referenceSNP <- fread('SNP/reference.1000G.maf.0.005.txt.gz', data.table = F)
```

For munging, all the Sum stats must have the rsIDs, all have rsID exept delange (2) GWAS.

Split the MarkerName column in ulccol_delange and crohn_delange into three different columns
```{r}
ulccol_delange <- separate(ulccol_delange, 'MarkerName', 
         sep= '_|_', 
         remove = T, 
         convert = F, 
         extra = 'warn', 
         into= c('chrPosition', 'Extra_1', 'Extra_2'))

crohn_delange <- separate(crohn_delange, 'MarkerName', 
         sep= '_|_', 
         remove = T, 
         convert = F, 
         extra = 'warn', 
         into= c('chrPosition', 'Extra_1', 'Extra_2'))
```

```{r inspect}
head(ulccol_delange)
head(crohn_delange)
```


```{r inspect}
dim(ulccol_delange) # 9588016      17
dim(crohn_delange) #9570787      17
```

Add the chrPosition to the reference dataset
```{r inspect}
head(referenceSNP)
dim(referenceSNP) # 9667224       6
```

Add the column 
```{r}
referenceSNP <- unite(referenceSNP, CHR, BP, remove = T, sep= ':', na.rm= F, col='chrPosition')
```

```{r inspect after insertion}
dim(referenceSNP) # 9667224       3
head(referenceSNP)
```

Create a reference file with only the columns of interest
```{r}
referenceSNP_short <- select(referenceSNP, SNP, A1, A2, chrPosition)
dim(referenceSNP_short) # 9667224       4
```

Match the columns with the reference and show the number of unmatched
```{r}
crohn_delange_rsID <- base::merge( x=crohn_delange, 
                             y=referenceSNP_short,
                             by.x= 'chrPosition',
                             by.y= 'chrPosition',
                             all.x= T,
                             all.y=F, 
                             sort=T,
                             )
sum(is.na(crohn_delange_rsID$SNP)) #
```

1 209 146 over 9 million of unmatched. 
The check if some NA SNPs are in the reference

```{r}
which(referenceSNP$`chrPosition` == '1:100000989') #should not be there and it is not there
which(referenceSNP$`chrPosition` == '1:569933') #should be there and it is there
```


Store the ones 
#crohn_delange_rsID_NA <- crohn_delange_rsID[is.na(crohn_delange_rsID$SNP) ,]
ulccol_delange_rsID <- base::merge( x=ulccol_delange, 
                                   y=referenceSNP_short,
                                   by.x= 'chrPosition',
                                   by.y= 'chrPosition',
                                   all.x= T,
                                   all.y=F, 
                                   sort=T
                                   )

#number of unmatched
sum(is.na(ulccol_delange_rsID$SNP)) #1 212 146


### match the columns with the reference for wightman
alzheimer_wightman <-unite(alzheimer_wightman, 
                           chr, PosGRCh37 ,
                           remove = F, 
                           sep= ':', 
                           na.rm= F, 
                           col='chrPosition')


alzheimer_wightman_rsID <- base::merge(x=alzheimer_wightman,
                                       y=referenceSNP_short, 
                                       by.x='chrPosition',
                                       by.y= 'chrPosition',
                                       all.x= T,
                                       all.y=F, 
                                       sort=T)
  
dim(alzheimer_wightman) # 12.688.339      8     
sum(is.na(alzheimer_wightman_rsID$SNP)) #3.584.435


### what happened to the ones without rsID
length(grep('rs', referenceSNP$SNP))
length(referenceSNP$SNP)
ulcolNA <- ulccol_delange_rsID[ which(is.na(ulccol_delange_rsID$SNP)), ]
head(ulccol_delange_rsID)
which(referenceSNP$`chr:Position` == '1:100000989')


#remove the NA columns from
crohn_delange_rsID_noNA <- crohn_delange_rsID[which(!is.na(crohn_delange_rsID$SNP)), ]
ulccol_delange_rsID_noNA <- ulccol_delange_rsID[which(!is.na(ulccol_delange_rsID$SNP)), ]
alzheimer_wightman_rsID_noNA <- alzheimer_wightman_rsID[ which(!is.na( alzheimer_wightman_rsID$SNP )), ] 


#create a fucntion that shows the p-values reported and calculated to see if is SE of logistic beta
is_se_logB <- function(BETA,SE, PVALUE) {
  p_calculated <- 2*pnorm((abs(BETA) / SE),lower.tail = F)
  p_reported <- PVALUE
  data.frame(p_calculated, p_reported)
}
#crohn_delange_rsID_noNA #std error of a logistic beta 
head(crohn_delange_rsID_noNA)                       
is_se_logB(crohn_delange_rsID_noNA$Effect, 
           crohn_delange_rsID_noNA$StdErr, 
           crohn_delange_rsID_noNA$P.value) #std error of a logistic beta 


#ulccol_delange_rsID_noNA std error of a logistic beta
is_se_logB(ulccol_delange_rsID_noNA$Effect,
           ulccol_delange_rsID_noNA$StdErr, 
           ulccol_delange_rsID_noNA$P.value ) #std error of a logistic beta

#alzheimer_kunkle_rsID
is_se_logB(alzheimer_kunkle_rsID$Beta,
           alzheimer_kunkle_rsID$SE, 
           alzheimer_kunkle_rsID$Pvalue) #std error of a logistic beta

#ra_okada confidence interval and not SE
#calculate the SE from confidence interval #SE = (upper limit – lower limit) / 3.92.
ra_okada_rsID <- separate(ra_okada_rsID,
                               'OR_95%CIup-OR_95%CIlow',
                               into = c('R_95%CIup', 'R_95%CIlow'), 
                               sep = '-', 
                               remove = F,
                               convert = T, 
                               extra = 'warn'
                                ) 

#asthma_demeanis_rsID
is_se_logB(asthma_demeanis_rsID$Multiancestry_beta_fix, 
           asthma_demeanis_rsID$Multiancestry_se_fix, 
           asthma_demeanis_rsID$Multiancestry_pval_fix) #standard error of logistic beta

#allergies_ferreira_rsID
is_se_logB(allergies_ferreira_rsID$BETA, 
           allergies_ferreira_rsID$SE, 
           allergies_ferreira_rsID$PVALUE) #standard error of logistic beta


#save the txt
write.table(crohn_delange_rsID_noNA, file = 'Outputs/Version1/V1_step1/crohn_delange_rsID_noNA.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(ulccol_delange_rsID_noNA, file = 'Outputs/Version1/V1_step1/ulccol_delange_rsID_noNA.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(allergies_ferreira, file ='Outputs/Version1/V1_step1/allergies_ferreira_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(alzheimer_kunkle, file = 'Outputs/Version1/V1_step1/alzheimer_kunkle_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(ssc_lopez, file= 'Outputs/Version1/V1_step1/ssc_lopez_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(ra_okada, file = 'Outputs/Version1/V1_step1/ra_okada_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(asthma_demeanis, file= 'Outputs/Version1/V1_step1/asthma_demeanis_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(alzheimer_wightman_rsID_noNA, file='Outputs/Version1/V1_step1/alzheimer_wightman_rsID_noNA.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )

#save the original with also the Non assigned rsiDs (NA)
write.table(crohn_delange_rsID, file = 'Outputs/Version1/V1_step1/crohn_delange_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(ulccol_delange_rsID, file = 'Outputs/Version1/V1_step1/ulccol_delange_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )
write.table(alzheimer_wightman_rsID, file='Outputs/Version1/V1_step1/alzheimer_wightman_rsID.txt',
            sep = "\t", quote = F, row.names = F, col.names=T )











